---
title: "Chronic 3,4-Methylenedioxymethamphetamine (MDMA) use is related to glutamate and GABA concentrations in the striatum, but not in the anterior cingulate cortex"
author: "Josua Zimmermann"
date: "21.4.2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning=F, message = F)
```


```{r}
rm(list=ls())

# load required packages
library(tidyverse)
library(knitr)
library(sjPlot)
library(ggsci)
library(patchwork)
library(robustbase)

source("used_functions.R")


# load data
data_path <- "Data"
df <- read.csv(file.path(data_path, "MRS_MDMA_data.csv"))

```


# RESULTS INCLUDED IN MAIN MANUSCRIPT


## LINEAR REGRESSION TO ESTIMATE GROUP DIFFERENCES

```{r}

# COMPUTE LINEAR REGRESSIONS TO ESTIMATE GROUP EFFECTS

# predictor of interest
predictor <- "Group"

# predictors to include as covariates in model
# - covariates regarding co-use of other substances are log-transformed (constant of 1 is added because of zero values)
covariates <- c("Age", "Gender", "log(Alcohol_amount_last_6months_gram + 1)", "log(THC_hairconcentration + 1)", "log(Stimulants_hairconcentration + 1)")

# measured regions
regions <- c("L Striatum", "ACC")

# measured outcomes of both sequences (MEGA-PRESS and PRESS)
outcomes <- c("GABA",  "GLX_MEGAdiff", "GLX_MEGAoff", "GLX_PRESS")


# initiate list to store model outputs
model_results <- list()

# loop over all regions and measures to compute linear regression
for (reg in regions){
  for (outc in outcomes){
    
    # get model formula
    model_formula <- as.formula(paste(paste0("scale(",outc,")"), paste(c(predictor, covariates), collapse = "+"), sep = "~"))
    # get model coefficients (and store in list)
    model_results[[reg]][[outc]] <- lm(model_formula, filter(df, Region == reg))    
    
  }
}


# SHOW TABLE WITH COMPLETE REGRESSION OUTCOMES

# labels for coefficients
pred_labels <- c("(Intercept)", "Group [MD]", "Age", "Gender [Male]", "Alcohol use", "Cannabis use", "Stimulant use")
# label for outcomes
dv_label <- c("GABA MEGAdiff", "GLX MEGAdiff", "GLX MEGAoff", "GLX PRESS")

# show table for the left striatum
tab_model(model_results[["L Striatum"]], pred.labels = pred_labels, dv.labels = dv_label, show.ci=F, title="Left Striatum", show.stat=T)

```

$~$
$~$

```{r}
# show table for the ACC
tab_model(model_results[["ACC"]], pred.labels = pred_labels, dv.labels = dv_label, show.ci=F, title="Anterior cingulate cortex", show.stat=T)

```

$~$
$~$

## BOXPLOTS TO SHOW GROUP DIFFERENCES

```{r}

# BOXPLOTS TO DISPLAY GROUP DIFFERENCES IN LEFT STIATUM
boxplot_GLX <- boxplot_groups("GLX_MEGAdiff", "Group", filter(df, Region=="L Striatum"), "GLX MEGAdiff [mol/kg]", c("Control", "MDMA"), "0.008")

boxplot_GABA <- boxplot_groups("GABA", "Group", filter(df, Region=="L Striatum"), "GABA MEGAdiff [mol/kg]", c("Control", "MDMA"), "0.31")


# PLOT CORRELATION BETWEEN GABA IN LEFT STRIATUM AND MDMA USE FREQUENCY
robcorr_GABA <- plot_robust_correlation("GABA", "MDMA_occasions_last_6months", filter(df, Region == "L Striatum" & Group == "MD"), "GABA MEGAdiff [mol/kg]", "log(MDMA occasions last 6 months)", TRUE, 5)


# plot all as subplots
boxplot_GLX + boxplot_GABA

robcorr_GABA

``` 


$~$
$~$

## EXPLORATORY ANALYSIS TO DISENTANGLE GLUTAMATE AND GLUTAMINE
To exploratory assess whether observed significant group difference in the left striatum is driven by glutamate or glutamine, we computed group differences for both metabolites.

```{r}

outcomes_expl <- c("GLU_MEGAdiff", "GLN_MEGAdiff")
region_expl <- "L Striatum"

# initiate list to store model outputs
model_results_expl <- list()

# loop over all regions and measures to compute linear regression
for (reg in region_expl){
  for (outc in outcomes_expl){
    
    # get model formula
    model_formula <- as.formula(paste(paste0("scale(",outc,")"), paste(c(predictor, covariates), collapse = "+"), sep = "~"))
    # get model coefficients (and store in list)
    model_results_expl[[reg]][[outc]] <- lm(model_formula, filter(df, Region == reg))    
    
  }
}


# SHOW TABLE WITH COMPLETE REGRESSION OUTCOMES

# show table for the left striatum
tab_model(model_results_expl[["L Striatum"]], pred.labels = pred_labels, dv.labels = c("Glutamate", "Glutamine"), show.ci=F, title="Left Striatum", show.stat=T)

```

$~$
$~$

## AGREEMENT OF PRESS AND MEGA-PRESS GLX MEASUREMENTS


```{r}


# PLOT CORRELATION BETWEEN GLX from MEGA-PRESS and PRESS

# left striatum
robcorr_GLXdiff_LS <- plot_robust_correlation("GLX_MEGAdiff", "GLX_PRESS", filter(df, Region == "L Striatum"),"GLX MEGAdiff [mol/kg]", "GLX PRESS [mol/kg]", FALSE, 1)

robcorr_GLXoff_LS <- plot_robust_correlation("GLX_MEGAoff", "GLX_PRESS", filter(df, Region == "L Striatum"), "GLX MEGAoff [mol/kg]", "GLX PRESS [mol/kg]", FALSE, 1)

robcorr_GLX_MEGA_LS <- plot_robust_correlation("GLX_MEGAdiff", "GLX_MEGAoff", filter(df, Region == "L Striatum"), "GLX MEGAdiff [mol/kg]", "GLX MEGAoff [mol/kg]", FALSE, 1)


# ACC
robcorr_GLXdiff_ACC <- plot_robust_correlation("GLX_MEGAdiff", "GLX_PRESS", filter(df, Region == "ACC"),"GLX MEGAdiff [mol/kg]", "GLX PRESS [mol/kg]", FALSE, 1)

robcorr_GLXoff_ACC <- plot_robust_correlation("GLX_MEGAoff", "GLX_PRESS", filter(df, Region == "ACC"), "GLX MEGAoff [mol/kg]", "GLX PRESS [mol/kg]", FALSE, 1)

robcorr_GLX_MEGA_ACC <- plot_robust_correlation("GLX_MEGAdiff", "GLX_MEGAoff", filter(df, Region == "ACC"), "GLX MEGAdiff [mol/kg]", "GLX MEGAoff [mol/kg]", FALSE, 1)

```


### Left striatum

```{r}

# Left striatum: print correlation of GLX between different sequences
robcorr_GLXdiff_LS + robcorr_GLXoff_LS

robcorr_GLX_MEGA_LS

```

$~$

### Anterior cingulate cortex

```{r}

# ACC: print correlation of GLX between different sequences
robcorr_GLXdiff_ACC + robcorr_GLXoff_ACC

robcorr_GLX_MEGA_ACC


```


$~$
$~$
$~$

# RESULTS INCLUDED IN THE SUPPLEMENT

## ROBUST REGRESSION TO ESTIMATE GROUP DIFFERENCES

To confirm reliablity of results, i.e., to ensure that outlier data points do not influence observed significant group differences, we compute robust regression models.

```{r}

# COMPUTE ROBUST REGRESSIONS TO ESTIMATE GROUP EFFECTS

# initiate list to store model outputs
model_results_robust <- list()

# loop over all regions and measures to compute linear regression
for (reg in regions){
  for (outc in outcomes){
    
    # get model formula
    model_formula <- as.formula(paste(paste0("scale(",outc,")"), paste(c(predictor, covariates), collapse = "+"), sep = "~"))
    # get model coefficients (and store in list)
    model_results_robust[[reg]][[outc]] <- lmrob(model_formula, filter(df, Region == reg))    
    
  }
}


# SHOW TABLE WITH COMPLETE "ROBUST" REGRESSION OUTCOMES

# show table for the left striatum
tab_model(model_results_robust[["L Striatum"]], pred.labels = pred_labels, dv.labels = dv_label, show.ci=F, title="Left Striatum", show.stat=T)

```

$~$
$~$

```{r}
# show table for the ACC
tab_model(model_results_robust[["ACC"]], pred.labels = pred_labels, dv.labels = dv_label, show.ci=F, title="Anterior cingulate cortex", show.stat=T)


```

$~$
$~$

## AGREEMENT OF PRESS and MEGA-PRESS REGARDING CREATINE AND NAA

```{r}

# correlation between creatine as well as NAA between PRESS and MEGA-PRESS

# left striatum
robcorr_CRE_LS <- plot_robust_correlation("CRE_MEGAoff", "CRE_PRESS", filter(df, Region == "L Striatum"),"Creatine MEGAoff [mol/kg]", "Creatine PRESS [mol/kg]", FALSE, 1)

robcorr_NAA_LS <- plot_robust_correlation("NAA_MEGAoff", "NAA_PRESS", filter(df, Region == "L Striatum"),"NAA MEGAoff [mol/kg]", "NAA PRESS [mol/kg]", FALSE, 1)


# ACC
robcorr_CRE_ACC <- plot_robust_correlation("CRE_MEGAoff", "CRE_PRESS", filter(df, Region == "ACC"),"Creatine MEGAoff [mol/kg]", "Creatine PRESS [mol/kg]", FALSE, 1)

robcorr_NAA_ACC <- plot_robust_correlation("NAA_MEGAoff", "NAA_PRESS", filter(df, Region == "ACC"),"NAA MEGAoff [mol/kg]", "NAA PRESS [mol/kg]", FALSE, 1)

```


### Left striatum

```{r}

# left striatum
robcorr_CRE_LS + robcorr_NAA_LS

```

$~$

# ACC

```{r}

# ACC
robcorr_CRE_ACC + robcorr_NAA_ACC

```
